---
title: "My Reproducible Report"
author: "Dean Marengi"
date: "`r Sys.Date()`"
output: html_document
  # html_document:
  #   theme: cerulean
---

## Overview

This file provides a quick look at how each component of an R Markdown file works together to produce an integrated report.

Remember, R Markdown allows us to:

-   **Combine narrative text, code, and output** into a single document
-   Incorporate mathematical expressions or equations
-   And render/export the .Rmd file to a variety of output file formats!
-   The YAML header specified in this `.Rmd` file enables us to render the output to:
    -   A Microsoft Word document (`.docx`)
    -   A `pdf` file
    -   Or `html` file

<br>

-   **Gapminder dataset/R package:** <https://cran.r-project.org/web/packages/gapminder/vignettes/gapminder.html>

## Load packages

In this code chunk, we are loading some relevant package libraries. By using the `include = FALSE` in our chunk options, we are telling the `.Rmd` file to exclude the code/output from the exported document.

```{r, include=FALSE}
# Packages
library(tidyverse)
library(gapminder)
library(kableExtra)

# Create new data frame object
data <- gapminder
```

## Create a plot in the below code chunk

Below we create a faceted figure using the gapminder dataset. We use the chunk options to specify the figure output height and width, as well as to create a figure caption. Notice that we can also use markdown syntax to format the caption text (e.g., **bold face text**)

```{r, fig.height=7, fig.width=10, fig.cap="**Life expectancy over time by country, stratified by continent.**", fig.align="center"}
# Create life expectancy by year figure
ggplot(data, aes(x = year, y = lifeExp, color = country)) +
  geom_line(show.legend = FALSE) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1950, 2010),
    breaks = seq(1950, 2010, 10)
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(20, 90),
    breaks = seq(20, 90, 10)
  ) +
  labs(x = "Year", y = "Life Expectancy at Birth (Years)") +
  facet_wrap(~ continent) +
  theme_bw() +
  theme(
    axis.text = element_text(face = "bold", color = "black"),
    axis.text.x = element_text(angle = 45, vjust = 0.6),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    panel.spacing.x = unit(1.5, "lines"),
    panel.spacing.y = unit(1.5, "lines")
  )
```

## How to embed a simple table (using the KableExtra package)

Here's a walkthrough of what each line of code is doing:

-   Reshape the `gapminder` dataset so it's in a long-data format
-   Filter for only life expectancy values obtained for the year 2007
-   Apply a group structure to the data frame (to compute grouped summary statistics)
-   Compute some simple summary statistics by continent
-   Round to two decimal places across numeric columns
-   Use the Kable package to generate a simple but nicely formatted table

```{r, R.options=list(width=50), warning=FALSE}
data %>%
  pivot_longer(lifeExp:gdpPercap, names_to = "parameters", values_to = "values") %>%
  filter(year == 2007 & parameters == "lifeExp") %>%
  group_by(continent) %>%
  summarize(Mean = mean(values, na.rm = TRUE),
            `Std. Dev` = sd(values, na.rm = TRUE),
            `Min.`= min(values, na.rm = TRUE),
            `Max.` = max(values, na.rm = TRUE)) %>%
  mutate(across(is.numeric, ~ round(., 2))) %>%
  kbl() %>%
  kable_classic()
```

### Some examples of how you can embed quantitative statements into your file with LaTeX.

-   Enclosing the expression in pairs of two dollar signs (`$$`..`$$`) generates a stand-alone equation. $$
    \begin{equation}
    \hat{Y}_i = \hat{\beta}_0 + \hat{\beta}_1 X_i + \hat{\epsilon}_i
    \end{equation}
    $$
-   In contrast, wrapping the expression in a pair of single dollar signs (`$`..`$`) creates an in-line quantitative expression.

<br>

-   This is some text with an in-line equation like $\hat{Y}_i = \hat{\beta}_0 + \hat{\beta}_1 X_i + \hat{\epsilon}_i$. Pretty cool, right?

<br> <br>
